Exploration

### 2018-03-10 exploration.Rmd
simulated data across time

```{r setup, include=F}
knitr::opts_chunk$set(echo = T)
pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
theme_set(theme_bw())
source("code/fn_IPM.R"); source("code/fn_aux.R")
lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean)) %>%
  mutate(x=as.integer(factor(.$left)),
         y=as.integer(factor(.$top,
                             levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 200
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 50  # ncells in IPM matrix
tmax <- 10
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
                    fl_x=c(1, -.01, -1, -.01, 0, .2), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.9, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=4, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,760])) + geom_tile()
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage & IPM matrix objects
p <- sp_i$p
setup_IPM_matrix(n, z.rng, buffer=0.25, discrete=1)
IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
d <- vector("list", length=n.cell)
B <- rpois(n.cell, n_i)
B_1 <- nSd <- D <- p_est.i <- rep(NA, n.cell)
sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd[,,1:2,] <- sdd.pr

# set up environmental covariates
n_sz <- length(p$s_z); n_sx <- length(p$s_x)
n_gz <- length(p$g_z); n_gx <- length(p$g_x)
n_flz <- length(p$fl_z); n_flx <- length(p$fl_x)
n_seedz <- length(p$seed_z); n_seedx <- length(p$seed_x)
X_s <- as.matrix(env.df[,1:n_sx])
X_g <- as.matrix(env.df[,1:n_gx])
X_fl <- as.matrix(env.df[,1:n_flx])
X_seed <- as.matrix(env.df[,1:n_seedx])

########
## loop through cells
########
# local growth
for(i in 1:n.cell) {
  d[[i]] <- data.frame(size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)

  #---- Truth: use assigned slopes to fill IPM matrix
  ### growth and survival
  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
  ### Fix eviction of individuals below/above size limits
  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
  ### fill P and F matrices
  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
  Ps[1,1,i] <- calc_staySB(p)
  ### store seed production
  Fb[z.i,z.i,i] <- outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                         X.seed=X_seed[i,], X.fl=X_fl[i,]) *h * p$p_est
  Fb[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                            X.seed=X_seed[i,], X.fl=X_fl[i,]) * h
  Fs[z.i,z.i,i] <- (1 - p$p_emig) * Fb[z.i,z.i,i]
  Fs[1,z.i,i] <- (1 - p$p_emig) * Fb[1,z.i,i]
  Fs[z.i,1,i] <- calc_rcrSB(y, p) * p$p_est

  #---- generate data
  ## survival
  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
  ## growth
  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
  ## flowering 
  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
  ## seed production
  E_seed <- d[[i]]$fl * 
    exp(size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  # d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
  nSd[i] <- sum(d[[i]]$seed, na.rm=T)
  ## seed bank: native
  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
    nSd[i] * (1 - p$p_emig) * (1 - p$rcr_dir) * p$s_SB
  ## recruits: native
  p_est.i[i] <- ifelse(p$NDD, sum(d[[i]]$surv==1)/nSd[i], p$p_est)
  rcr.n <- ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p_est.i[i] +
                   B[i] * p$rcr_SB * p_est.i[i]) # ≥ 2 for IPMpack sd(rcr_z')
  rcr.i <- (1:rcr.n) + n_i[i]
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
  ## emigrants
  sdd[,,3,i] <- nSd[i] * p$p_emig * sdd[,,1,i]
}
# dispersal
for(i in 1:n.cell) {
  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes

  #---- Truth: add immigrants
  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)
  p.ij <- sdd[,,1,][sdd.j]
  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
    Reduce(`+`, map2(sdd.j[,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[,3]] %*% (p$p_emig*p.ij)

  #---- generate data: add immigrants
  # total immigrants
  D[i] <- sum(sdd[,,3,][sdd.j])
  # seed bank: immigrants
  B_1[i] <- B_1[i] + D[i] * (1 - p$rcr_dir) * p$s_SB
  # recruits: immigrants
  rcr.imm <- ceiling(D[i] * p$rcr_dir * p_est.i[i])
  rcr.i <- (1:rcr.imm) + nrow(d[[i]])
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
}
# density dependence
if(p$NDD) {
  lam.t <- p_est.t <- matrix(nrow=tmax-1, ncol=n.cell)
  p_est.t[1,] <- p$p_est
  for(i in 1:n.cell) {
    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i)
    for(k in 1:(tmax-1)) {
      F.t <- Fs[,,i]
      p_est.t[k,i] <- min(195/(sum(F.t[z.i,] * Nt[z.i,,i,k])/p$p_est), p$p_est)
      F.t[z.i,] <- F.t[z.i,]/p$p_est * p_est.t[k,i]
      IPM.t <- Ps[,,i] + F.t
      Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
      lam.t[k,i] <- Re(eigen(IPM.t)$values[1])
    }
  }
}
IPMs <- Ps + Fs
lam.df <- env.df
lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D) %>%
  mutate(nSdStay=nSd*(1-p$p_emig), nSdLeave=nSd*p$p_emig) %>%
  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.tmax))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSd)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=B1>B0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=(D-nSdStay)/nSdStay)) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay < D & lambda < 1)) +
  geom_tile()
par(mfrow=c(2,2))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOth, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)

>>>>

```{r setup, include=F}
knitr::opts_chunk$set(echo = T)
pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
suppressMessages(invisible(lapply(pkgs, library, character.only=T)))
theme_set(theme_bw())
source("code/fn_IPM.R"); source("code/fn_aux.R"); source("code/fn_sim.R")
lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean)) %>%
  mutate(x=as.integer(factor(.$left)),
         y=as.integer(factor(.$top,
                             levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 100
regPlots <- F
ipmPlots <- F
nPlots <- 5
z.rng <- c(1, 12)
n <- 10  # ncells in IPM matrix
tmax <- 30
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=100, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,740])) + geom_tile()
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage & IPM matrix objects
p <- sp_i$p
setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
d <- sdd.j <- vector("list", length=n.cell)
B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
nSd <- D <- p_est.i <- matrix(nrow=n.cell, ncol=tmax)
lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd[,,1:2,] <- sdd.pr

# set up environmental covariates
n_z <- list(s=length(p$s_z), g=length(p$g_z),
            fl=length(p$fl_z), seed=length(p$seed_z))
n_x <- list(s=length(p$s_x), g=length(p$g_x),
            fl=length(p$fl_x), seed=length(p$seed_x))
X <- list(s=as.matrix(env.df[,1:n_x$s]), g=as.matrix(env.df[,1:n_x$g]),
          fl=as.matrix(env.df[,1:n_x$fl]), seed=as.matrix(env.df[,1:n_x$seed]))

########
## loop through cells
########

#---- Truth: use assigned slopes to fill IPM matrix
## local growth
for(i in 1:n.cell) {
  Ps[,,i] <- fill_P(h, y, z.i, n, p, n_z, n_x, X$s[i,], X$g[i,])
  Fb[,,i] <- fill_F(h, y, z.i, n, p, n_z, n_x, X$fl[i,], X$seed[i,])
  Fs[z.i,z.i,i] <- (1-p$p_emig) * Fb[z.i,z.i,i]
  Fs[1,z.i,i] <- (1-p$p_emig) * Fb[1,z.i,i]
  Fs[z.i,1,i] <- Fb[z.i,1,i]
}
## dispersal & density dependence
for(i in 1:n.cell) {
  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
  sdd.j[[i]] <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE) # j cells with target i
  p.ij <- sdd[,,1,][sdd.j[[i]]] # p(j to i | immigrant seed from j)
  Fs[z.i,z.i,i] <- Fs[z.i,z.i,i] + 
    Reduce(`+`, map2(sdd.j[[i]][,3], p$p_emig*p.ij, ~(Fb[z.i,z.i,.x] * .y)))
  Fs[1,z.i,i] <- Fs[1,z.i,i] + Fb[1,z.i,sdd.j[[i]][,3]] %*% (p$p_emig*p.ij)
  if(p$NDD) {
    Nt[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
    F.i <- F.t <- Fs[,,i]/p$p_est  # F.s[,,i] was already multiplied by p$p_est
    for(k in 1:(tmax-1)) {
      p_est.t[i,k] <- min(p$NDD_n/(sum(F.i[z.i,]*Nt[z.i,,i,k])), p$p_est)
      F.t[z.i,] <- F.i[z.i,] * p_est.t[i,k]
      IPM.t <- Ps[,,i] + F.t
      Nt[,,i,k+1] <- IPM.t %*% Nt[,,i,k]
      lam.t[i,k] <- Re(eigen(IPM.t)$values[1])
    }
  }
}
IPMs <- Ps + Fs

#---- Realization: generate simulated data
## local growth
for(i in 1:n.cell) {
  d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
  E <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E, p, lo, hi)
  nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
  sdd[,,3,i] <- nSd[i,1] * p$p_emig * sdd[,,1,i]
}
## dispersal & density dependence
for(i in 1:n.cell) {
  D[i,1] <- sum(sdd[,,3,][sdd.j[[i]]])
  p_est.i[i,1] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,1]+D[i,1]),p$p_est), p$p_est)
  d[[i]] <- sim_recruits(d[[i]], p_est.i[i,1], nSd[i,1], B[i,1], D[i,1], p)
  B[i,2] <- sim_seedbank(nSd[i,1], B[i,1], D[i,1], p)
}
## iterate forward
for(k in 2:tmax) {
  for(i in 1:n.cell) {
    size.k <- d[[i]]$sizeNext[!is.na(d[[i]]$sizeNext) & d[[i]]$yr == k-1]
    if(length(size.k) > 0) {
      k.index <- (1:length(size.k)) + nrow(d[[i]])
      d[[i]][k.index, "size"] <- size.k
      d[[i]][k.index, "yr"] <- k
      E <- sim_expected(d[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
      d[[i]] <- sim_realized(d[[i]], k.index, E, p, lo, hi)
      nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
    } else {
      nSd[i,k] <- 0
    }
    sdd[,,3,i] <- nSd[i,k] * p$p_emig * sdd[,,1,i]
  }
  for(i in 1:n.cell) {
    D[i,k] <- sum(sdd[,,3,][sdd.j[[i]]])
    p_est.i[i,k] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,k]+D[i,k]),p$p_est), p$p_est)
    d[[i]] <- sim_recruits(d[[i]], p_est.i[i,k], nSd[i,k], B[i,k], D[i,k], p)
    B[i,k+1] <- sim_seedbank(nSd[i,k], B[i,k], D[i,k], p)
  }
}

lam.df <- env.df
lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(nSeed=nSd[,tmax], B0=B[,1], Btmax=B[,tmax+1], D=D[,tmax], 
             N=map_dbl(d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             N.surv=map_dbl(d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             N.rcr=map_dbl(d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
  add_column(Ntmax=apply(Nt[,,,tmax],3,sum), lam.tmax=lam.t[tmax-1,])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.tmax))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Ntmax)*(Ntmax>10))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.tmax>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=Ntmax>1 & lam.tmax<1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.surv))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.rcr))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay/D)) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=1)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay < D & lambda < 1)) +
  geom_tile()
par(mfrow=c(2,2))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)

sample.cells <- sample.int(n.cell, nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
  mx <- cbind(1, as.matrix(xx))
  xxx <- seq(0, hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(mx[,2], antilogit(mx[,1:n_sz]%*%p$s_z + c(X_s[i,]%*%p$s_x)), col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(mx[,2], mx[,1:n_gz]%*%p$g_z + c(X_g[i,]%*%p$g_x), col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(mx[,2], antilogit(mx[,1:n_flz]%*%p$fl_z + c(X_fl[i,]%*%p$fl_x)),
            col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(mx[,2], exp(mx[,1:n_seedz]%*%p$seed_z + c(X_seed[i,]%*%p$seed_x)),
            col=2)
    plot(y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, y)
  for(i in sample.cells) {
    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
    z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
    image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
      points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
    image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
    image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
               xlab="Size (t)", ylab="Size (t+1)")
  }
}
image.plot(x, x, t(apply(Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, t(apply(Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, sens(t(apply(IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
```
